---
title: "REC_R"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
sapply(list.files("user_functions/", pattern = ".R$", full.names = TRUE), source)
```

This project translates algorithms implemented by **Rate Estimation from Concentration (REC)** in matlab code into R code. Some minor changes to the workflow are also implemented. 

REC was originally developed by Karsten Lettmann ([Lettmann et al. 2012](https://www.sciencedirect.com/science/article/abs/pii/S0272771411000229)) and can be downloaded [here](https://uol.de/icbm/physikalische-ozeanographie-theorie/downloads).

## Preparing data
The matlab version requires data input from a 'setup', a folder containing individual data files named \*_C.txt, \*_phi.txt, \*_omega.txt , \*_beta.txt, \*_D.txt, and \*_Db.txt, all files contain z (depth) values. This R version take as input a data frame with columns `c(z, C, phi, omega, beta, D, Db)`. Setups can still be used and data imported from them using the function `import_from_setup()`. `import_from_setup()` takes as argument the path to a setup folder, finds the required files in the setup folder and merges them based on z (depth) values, but ignores all other files so that results (or other) files could be saved in the setup without interfering with the import. Columns z, C, phi and D have to be supplied and cannot be all zeros; columns omega, beta, and Db may not be supplied and will the be substituted with all zeros.
*C* denotes concentration of the water constituent. &phi; (phi) denotes the sediment porosity, &omega; (omega) the sedimentation rate combined with pore water advection, *D* the molecular diffusion coefficient (corrected for sediment properties), *D~B~* (Db) denotes the bioturbation coefficient, &beta; (beta) the coefficient of irrigation.

Here we import data from the original example setup.

```{r}
df <- import_from_setup("test_case_2_data_delta/")
head(df)
```

## Running the main function
The example data is passed to the main function `rec()` as data frame. `rec()` takes a number of arguments to guide the modeling of the concentration profile and rates at a supplied depth resolution. The function then interpolates input data to the required resolution, passes them together with input arguments to a calculation function `calculate_con_rates_lin_sys_Tikhonov_mean_rate_2()`, and compiles the results for output as a names list. Internally, `calculate_con_rates_lin_sys_Tikhonov_mean_rate_2()` calls a few other functions.

```{r}
test <- rec(original_data = df,
            N_c = 101,               # Number of computational grid points
            C_water = 25e3,          # Nutrient concentration in water column (only important for irrigation)
            # parameters for Tikhonov regularization
            lambda = 1,              # 'smoothing' parameter lambda
            alpha_min = 8,           # lowest alpha value for Tikhonov regularisation and ratio criterion ( actually log_10(alpha_min) )
            alpha_max = 15,          # largest alpha value for Tikhonov regularisation and ratio criterion ( actually log_10(alpha_max) )
            N_alpha = 301,           # Number of ratio criterion evaluations in the alpha interval, to find the minimum
            # setting the boundary conditions for the nutrient concentration
            bnd_cond_type_z_min = 1, # type of boundary condition at the top: 1: for concentration / 2: for derivative
            bnd_cond_C_z_min = 25e3, # value of nutrient concentration or derivative at top
            bnd_cond_type_z_max = 1, # type of boundary condition at the bottom: 1: for concentration / 2: for derivative
            bnd_cond_C_z_max = 5e3   # value of nutrient concentration or derivative at bottom
            )
```

```{r}
summary(test)
```
The list contains the *input data* as a data frame; the *input parameters* entered as arguments of `rec()` as a list; the interpolated data (i.e., the input data at a depth resolution passed to  `rec()` via `N_c`, minus the z column) as a data frame; the output data of z in the required resolution and the corresponding *interpolated concentrations* and *calculated rates* as a data frame; the *optimum alpha* value; a vector of all tested alpha values and a vector of corresponding Tikhonov criterion values.

## Visualize results
The function `plot_rec()` makes it easy to plot results for a quick overview. It works on the output from `rec()`, so all plots (and more) can also be produced directly from the output.

By default `plot_rec()` plots a two panel depth profile of input and interpolated concentrations and calculated rates. The argument `type` can be used to produce figures of the input data or of the Tikhonov criterion as a function of alpha. The latter is useful to check if a local minimum in the Tikhonov criterion was actually found and a corresponding optimal alpha value used. This optimal alpha is shown on the figure as a green dot. Printed output from `rec()` will also warn if no local minimum was found and the absolute minimum is used as the location of the optimal alpha.

```{r}
plot_rec(test)
```

```{r}
plot_rec(test, type = "localmin")
```

```{r}
plot_rec(test, type = "input")
```

## Calculate boundary fluxes
Fluxes of the constituent of "C" across the upper and lower boundary of the depth profile can be calculated with `boundary_fluxes()`. This is useful e.g. in studies of benthic-pelagic coupling. The function accepts the output of `rec()`. By default, an explanation of the calculation is printed to the console, this can be suppressed by setting the argument `explain = FALSE`.

```{r}
boundary_fluxes(test)
```

## Calculate integrated rates
The rates of production of a constituent for "C" can be integrated using `integrate_rates()`. The integration interval can be defined using the `z_min` and `z_max` arguments but if no values for these are supplied, rates will be integrated over the entire z range. By default, the calculation will be explaind as text printed to the console but this can be suppresed by setting the argument `explain = FALSE`.

```{r echo=FALSE}
plot(test$output_data$z~test$output_data$rate, 
     ylim = rev(range(test$output_data$z)), ylab = "z", xlab = "rate", type = "l", col = "red")
abline(h = c(0,30), col = "blue")
```

```{r}
integrate_rates(test)
integrate_rates(test, z_min = 0, z_max = 30, explain = FALSE)
```
